Evolution of pre-copulatory and post-copulatory inbreeding behaviour and associated polyandry

### 2016-11-11 pre_post_biorxiv.tex
Several tweaks, mostly fix typos

Inbreeding often greatly reduces the fitness of inbred offspring \cite[termed `inbreeding depression';][]{Charlesworth1999, Keller2002, Charlesworth2009}. Such strong inbreeding depression is widely hypothesised to drive evolution of inbreeding avoidance, which can be enacted through multiple reproductive decisions in species with obligate biparental reproduction \cite[][]{Parker1979, Parker2006, Pusey1996, Szulkin2012}. From the perspective of a focal breeding female, inbreeding avoidance might be achieved by avoiding mating with related males (i.e., pre-copulatory inbreeding avoidance), or by biasing fertilisation towards unrelated males after mating has occurred, meaning that related males are less likely to sire offspring (i.e., post-copulatory inbreeding avoidance). Importantly, evolution of post-copulatory inbreeding avoidance requires that females are polyandrous, defined as mating with multiple males during a single reproductive bout. Further, such polyandry could itself evolve because it allows females to mate with additional unrelated males following initial mating with a relative, thereby facilitating pre-copulatory inbreeding avoidance even in the absence of any post-copulatory female choice or otherwise biased fertilisation (e.g., given a 'fair raffle' among sperm provided by females' mates). As a reproductive strategy, polyandry can thereby simultaneously allow females to mate with less closely related males and creates the opportunity for further inbreeding avoidance enacted through active post-copulatory choice. Polyandry has consequently been widely hypothesised to evolve as an adaptation to allow females to avoid inbreeding and thereby increase offspring fitness \cite[][]{Zeh1997, Jennions2000, Tregenza2002, Akcay2007}. However, despite such well-established hypotheses, no modelling investigates conditions under which pre-copulatory or post-copulatory inbreeding avoidance, or both strategies, are likely to evolve, or predicts how such evolution will feed back to affect underlying evolution of polyandry.

Meanwhile, empirical studies of diverse species have found evidence for female inbreeding avoidance enacted through pre-copulatory or post-copulatory routes, or both. For example, pre-copulatory inbreeding avoidance has been documented in weevils, birds, and butterflies \cite[e.g.,][]{Kuriwada2011, Kingma2013, Fischer2015}, and post-copulatory inbreeding avoidance has been documented in crickets \cite[e.g.,][]{Simmons2006, Bretman2009}. Evidence of both pre-copulatory and post-copulatory inbreeding avoidance has accumulated for some well-studied species, including Trinidadian guppies \cite[\textit{Poecilia reticulata};][]{Gasparini2011, Daniel2015} and house mice \cite[\textit{Mus domesticus};][]{Potts1991, Firman2015}. While most studies consider either pre-copulatory or post-copulatory inbreeding avoidance in isolation, a small number of studies consider both. Accordingly, \cite{Liu2014} found evidence of pre-copulatory, but not post-copulatory, inbreeding avoidance in cabbage beetles (\textit{Colaphellus bowringi}). Other studies found no evidence of pre-copulatory or post-copulatory inbreeding avoidance despite strong inbreeding depression \cite[e.g.,][]{Reid2014, Reid2015}. These empirical studies demonstrate that diverse combinations of pre-copulatory and post-copulatory inbreeding avoidance occur in nature. However, there is as yet no modelling framework that predicts \textit{a priori} what combinations should be favoured by selection when both pre-copulatory and post-copulatory inbreeding avoidance, and polyandry, can evolve. Such a framework would allow prediction of the form of mating system evolution across different systems that experience inbreeding depression and allow existing empirical studies to be compared to model predictions.

@@

Previous models implicitly or explicitly considered the fate of a rare allele underlying pre-copulatory inbreeding avoidance in a population initially fixed for random mating \cite[e.g.,][]{Parker1979, Parker2006, Duthie, Duthie2016a}. Such models thereby isolate the invasion fitness of this single strategy. However, when both pre-copulatory and post-copulatory strategies can affect realised inbreeding, it cannot be assumed that both strategies will invade simultaneously, nor that invasion fitness of one strategy will be independent of the other For example, if pre-adaptation or rapid selection results in the fixation of alleles underlying adaptive pre-copulatory inbreeding avoidance, then alleles underlying post-copulatory inbreeding avoidance might be precluded from invading a population in which post-copulatory choice is not already expressed because the effect of such invading alleles could be weak or negligible.

@@

\section*{Model}

We model evolution of polyandry and pre-copulatory and post-copulatory inbreeding avoidance by tracking reproductive interactions among individual females and males in a small focal population. Individuals are defined as related if they share a common ancestor within the population pedigree (i.e., non-zero kinship). Each individual has 10 diploid loci, and therefore 20 alleles, underlying each of the three reproductive strategy traits: tendency for polyandry ($P_{a}$), pre-copulatory inbreeding avoidance ($M_{a}$, i.e., `mating' alleles), and post-copulatory inbreeding avoidance ($F_{a}$ i.e., `fertilisation' alleles). Alleles can take any real number \cite[i.e., a continuum-of-alleles model;][]{Kimura1965, Lande1976, Reeve2000, Bocedi2014} and combine additively to affect genotypes ($P_{g}$, $M_{g}$, and $F_{g}$) and phenotypes ($P_{p}$, $M_{p}$, and $F_{p}$) for each trait. In overview, each generation proceeds with females paying costs, expressing polyandry, mating, and fertilisation. Offspring inherit a randomly sampled allele from each parent at each locas with no physical linkage. Alleles can then mutate and offspring express inbreeding depression. Immigrants arrive in the population and density regulation limits population growth. Key individual traits, parameter values, and variables are described in Table 1.

@@

After costs are paid, each remaining female chooses $N_{males}$ males to mate with, where $N_{males}$ is calculated by sampling from a Poisson distribution such that $N_{males} = Poisson(P_{p}) + 1$. This ensures that all surviving females choose at least one mate, and causes some stochastic variation aroudn the expected mean $N_{males}$ of $P_{p}+1$. A female's value of $P_{p}$ depends on the sum of her 20 $P_{a}$ allele values, which might be negative or positive. However, because females cannot mate with a negative number of males, $P_{p}$ is constrained to be zero or positive such that $P_{p} = 0$ when $P_{g} < 0$, but $P_{p} = P_{g}$ when $P_{g} \geq 0$. Polyandry is therefore modelled as a threshold trait \cite[][]{Lynch1998, Roff1996, Roff1998, Duthie} that is influenced by continuous genetic variation but only expressed at the threshold value of $P_{g} > 0$. Any negative $P_{g}$ value therefore generates phenotypic monandry, while positive $P_{g}$ values can lead to different degrees of realised polyandry (i.e., different numbers of mates, $N_{males}$).

@@

The viability of a focal female $i$'s offspring ($\Psi_{\textrm{off}}$) decreases as a log-linear function of her kinship with the sire of her offspring $j$ ($k_{ij}$) and inbreeding depression slope $\beta$,
\begin{equation}
\Psi_{\textrm{off}} = e^{-\beta k_{ij}}
\end{equation}
Here, $\beta$ models the number of haploid lethal equivalents that exist as deleterious recessive alleles in the gametes of $i$ and $j$, and which might be homozygous in offspring and reduce viability. Equation 1 assumes independent allelic effects, generating multiplicative effects on offspring viability \cite[][]{Morton1956, Mills1994}. We model inbreeding depression as having an absolute rather than relative effect on offspring viability) i.e., hard rather than soft selection), so that the effect of $\beta$ is consistent across generations and different parameter combinations. We also assume that inbreeding always decreases offspring viability, $\beta > 0$ (i.e., no outbreeding depression). Therefore, because $k_{ij}$ is constrained to values between zero and one, $-\beta \times k_{ij} \leq 0$. Values of $\Psi_{\textrm{off}}$ must therefore be between zero (if $-\beta \times k_{ij}$ is very negative) and one (if $-\beta \times k_{ij} = 0$). We therefore define $\Psi_{\textrm{off}}$ as the probability that an offspring is viable, and sample its realised viability (versus mortality) using a Bernoulli trial.

@@

\subsection*{Model analysis}

@@

When post-copulatory inbreeding avoidance alleles ($F_{a}$) were fixed to zero and could not evolve, $P_{a}$ values were always negative (Figure 1A,C,E,G). This shows that there is selection against unconditional polyandry even given zero direct cost ($c_{P}=0$; Figure 1A). This is because polyandrous females sampled more males from the available population and were consequently likely to mate with some relatives and hence produce some inbred offspring with low viability \cite[][]{Duthie}.

When post-copulatory inbreeding avoidance was allowed to evolve, mean $P_{a}$ values were substantially higher than in comparable simulations where post-copulatory inbreeding avoidance could not evolve (Figure 1B,D,F,H). Allowing evolution of post-copulatory inbreeding avoidance alongside pre-copulatory inbreeding avoidance therefore facilitated evolution of polyandry. However, $P_{a}$ values were expeted to be positive only given a low cost of polyandry ($c_{P} < 0.005$, Figure 1A,B). This implies that, even given strong inbreeding depression in offspring viability, inbreeding avoidance might generally be a weak force driving polyandry evolution.

Strong post-copulatory inbreeding avoidance consistently evolved in all simulations where such such evolution was allowed (Figure 1B,D,F,H), even when $P_{a}$ values were expected to be slightly negative (i.e., when there is selection against polyandry, Figure 1F,H). This reflects the threshold nature of phenotypic expression of polygenic polyandry, wherein random sampling of alleles means that polyandry is commonly expressed (i.e., $P_{g}>0$) even when mean $P_{a}$ values are negative (Figure 2). This means that there is commonly opportunity for expression of post-copulatory inbreeding avoidance and associated selection that drives evolution of $F_{a}$ allele values.

@@

 {\color{Gray}
\begin{adjustwidth}{-2.2in}{0in}
\mysidecaption{0.68}
{%
   \includegraphics[height=18cm]{figures/poly_costs.pdf}%
}%
{%
   \captionof{figures/poly_costs.pdf}\begin{justify}\vspace{10 mm} \textbf{Figure 1:} Mean allele values underlying tendency for polyandry (red), pre-copulatory inbreeding strategy (blue), and post-copulatory inbreeding strategy (black) from simulations where post-copulatory inbreeding strategy is (A, C, E, and G) fiext to zero (i.e., random fertilisation) or (B, D, F, H) allowed to evolve freeely. Costs of polyandry ($c_{p}$) increase across rows from zero (A and B) to 0.01 (G and H). Mean allele values (solid lines) and associated standard errors (shading) are calculated across all individuals within a population over 40000 generations across 40 replicate populations. Negative mean allele values indicate inbreeding avoidance or tendency for monandry, and positive values indicate inbreeding preference or tendency for polyandry. Dotted lines demarcate mean allele values of zero. \end{justify}{\t}%
}[t]
\end{adjustwidth}
}

@@

The opportunity to adjust inbreeding has been widely considered as a driver of adaptive evolution of polyandry \cite[][]{Tregenza2002, Foerster2003, Akcay2007, Varian-Ramos2012, Kingma2013, Lehtonen2015, Reid2014}. Recently, \cite{Duthie} found the conditions under which selection for polyandry and pre-copulatory inbreeding avoidance is predicted to be highly restrictive. However, \cite{Duthie} assumed that fertilisation of offspring was unbiased with respect to mate kinship, meaning that females could not express post-copulatory choice. In contrast, we found that when post-copulatory inbreeding avoidance could evolve, selection for polyandry was greatly strengthened (Figure 1). The proposition that polyandry might facilitate cryptic female choice among males of varying compatibility is not new \cite[e.g.,][]{Zeh1997}, but the existence of selection for post-copulatory inbreeding avoidance causing feedback to select for polyandry has not been formally modelled. Our model therefore has widespread implications for empirical studies of evolution of polyandry.

@@

Recent studies have shown how females can successfully bias fertilisation after mating has occurred. For example, when female guppies (*Poecilia reticulata*) were artificially inseminated with equal quantities of sperm from full-siblings and unrelated males, more eggs were fertilised by unrelated males because the velocities of full siblings' sperm were reduced by females' ovarian fluids \cite[][]{Gasparini2011}. In black field crickets (\textit{Teleogryllus commodus}), females attempt to remove the spermatophores of unwanted males after copulation, and are capable of controlling sperm transfer to spermatheca after copulation occurs \cite{Bussiere2006, Tuni2013}. Our results suggest that evolution of such post-copulatory inbreeding avoidance might increase selection for further polyandry, especially if pre-copulatory inbreeding avoidance is costly (Figure 1).

@@

In our model, one cost free inbreeding strategy precluded another more costly inbreeding strategy from persisting in a focal population (Figure 2), meaning that long-term persistence of both pre-copulatory and post-copulatory inbreeding avoidance might not be expected in populations given cost asymmetry. Direct costs therefore strongly affected the evolution of the costly phenotype(s), meaning that quantifying costs is necessary for predicting evolution of phenotypes associated with inbreeding strategies. Quantifying direct costs of phenotypes associated with polyandry and mate choice is empirically challenging \cite[][]{Pomiankowski1987, Kokko2003, Reid2015}. \cite{Pomiankowski1987} categorised four types of costs that are relevant to mate choice, which we suggest are also relevant costs for polyandry, and include costs of elevated risks associated with (1) predation or (2) disease transmission, and costs incurred through (3) loss of time or (4) depletion of energy. In our model, we interpreted a cost of polyandy as an elevated risk of predation as might be incurred while searching for or courting mates \cite[e.g.,][]{Rowe1988, Rowe1994}, a cost of pre-copulatory inbreeding strategy as a lost of time \cite[i.e., risk of not finding a mate in time due to choosiness; e.g.,][]{Kokko2013}, and a cost of post-copulatory inbreeding strategy as depletion of energy. While these costs are biologically realistic, different costs will be relevant for different populations and be realised in different ways. For example, increased risk of disease transmission has been observed for highly polyandrous females \cite[][]{Roberts2015a}, but such a cost would more realistically be applied to the number of realised mates a female has rather than her tendency for engaging in polyandry. Future studies will therefore need to carefully consider the magnitudes and realisation of direct costs to accurately predict evolution of inbreeding strategies in populations.

>>>>

Inbreeding often greatly reduces the fitness of inbred offspring \cite[termed `inbreeding depression';][]{Charlesworth1999, Keller2002, Charlesworth2009}. Such strong inbreeding depression is widely hypothesised to drive evolution of inbreeding avoidance, which can be enacted through multiple reproductive decisions in species with obligate biparental reproduction \cite[][]{Parker1979, Parker2006, Pusey1996, Szulkin2012}. From the perspective of a focal breeding female, inbreeding avoidance might be achieved by avoiding mating with related males (i.e., pre-copulatory inbreeding avoidance), or by biasing fertilisation towards unrelated males after mating has occurred, meaning that related males are less likely to sire offspring (i.e., post-copulatory inbreeding avoidance). Importantly, evolution of post-copulatory inbreeding avoidance requires that females are polyandrous, defined as mating with multiple males during a single reproductive bout. Further, such polyandry could itself evolve because it allows females to mate with additional unrelated males following initial mating with a relative, thereby facilitating pre-copulatory inbreeding avoidance even in the absence of any post-copulatory female choice or otherwise biased fertilisation (e.g., given a `fair raffle' among sperm provided by females' mates). As a reproductive strategy, polyandry can thereby simultaneously allow females to mate with less closely related males and create the opportunity for further inbreeding avoidance enacted through active post-copulatory choice. Polyandry has consequently been widely hypothesised to evolve as an adaptation to allow females to avoid inbreeding and thereby increase offspring fitness \cite[][]{Zeh1997, Jennions2000, Tregenza2002, Akcay2007}. However, despite such well-established hypotheses, no modelling investigates conditions under which pre-copulatory or post-copulatory inbreeding avoidance, or both strategies, are likely to evolve, or predicts how such evolution will feed back to affect underlying evolution of polyandry.

Meanwhile, studies of diverse species have found evidence for female inbreeding avoidance enacted through pre-copulatory or post-copulatory routes, or both. For example, pre-copulatory inbreeding avoidance has been documented in weevils, birds, and butterflies \cite[e.g.,][]{Kuriwada2011, Kingma2013, Fischer2015}, and post-copulatory inbreeding avoidance has been documented in crickets \cite[e.g.,][]{Simmons2006, Bretman2009}. Evidence of both pre-copulatory and post-copulatory inbreeding avoidance has accumulated for some well-studied species, including Trinidadian guppies \cite[\textit{Poecilia reticulata};][]{Gasparini2011, Daniel2015} and house mice \cite[\textit{Mus domesticus};][]{Potts1991, Firman2015}. While most studies consider either pre-copulatory or post-copulatory inbreeding avoidance in isolation, a small number of studies consider both. Accordingly, \cite{Liu2014} found evidence of pre-copulatory, but not post-copulatory, inbreeding avoidance in cabbage beetles (\textit{Colaphellus bowringi}). Other studies found no evidence of pre-copulatory or post-copulatory inbreeding avoidance despite strong inbreeding depression \cite[e.g.,][]{Reid2014, Reid2015}. These studies demonstrate that diverse combinations of pre-copulatory and post-copulatory inbreeding avoidance occur in nature. However, there is as yet no modelling framework that predicts what combinations should be favoured by selection when both pre-copulatory and post-copulatory inbreeding avoidance, and polyandry, can evolve. Such a framework would allow prediction of the form of mating system evolution across different systems that experience inbreeding depression and allow existing empirical studies to be compared to model predictions.

\subsection*{Verbal theory}

@@

Previous models implicitly or explicitly considered the fate of a rare allele underlying pre-copulatory inbreeding avoidance in a population initially fixed for random mating \cite[e.g.,][]{Parker1979, Parker2006, Duthie, Duthie2016a}. Such models thereby isolate the invasion fitness of this single strategy. However, when both pre-copulatory and post-copulatory strategies can affect realised inbreeding, it cannot be assumed that both strategies will invade simultaneously, nor that invasion fitness of one strategy will be independent of the other. For example, if pre-adaptation or rapid selection results in the fixation of alleles underlying adaptive pre-copulatory inbreeding avoidance, then alleles underlying post-copulatory inbreeding avoidance might be precluded from invading a population in which post-copulatory choice is not already expressed because the effect of such invading alleles could be weak or negligible.

@@

\section*{Methods}

We model evolution of polyandry and pre-copulatory and post-copulatory inbreeding avoidance by tracking reproductive interactions among individual females and males in a small focal population. Individuals are defined as related if they share a common ancestor within the population pedigree (i.e., non-zero kinship). Each individual has 10 diploid loci, and therefore 20 alleles, underlying each of the three reproductive strategy traits: tendency for polyandry ($P_{a}$), pre-copulatory inbreeding avoidance ($M_{a}$, i.e., `mating' alleles), and post-copulatory inbreeding avoidance ($F_{a}$, i.e., `fertilisation' alleles). Alleles can take any real number \cite[i.e., a continuum-of-alleles model;][]{Kimura1965, Lande1976, Reeve2000, Bocedi2014} and combine additively to affect genotypes ($P_{g}$, $M_{g}$, and $F_{g}$) and phenotypes ($P_{p}$, $M_{p}$, and $F_{p}$) for each trait. In overview, each generation proceeds with females paying costs, expressing polyandry, mating, and fertilisation. Offspring inherit a randomly sampled allele from each parent at each locus with no physical linkage. Alleles can then mutate and offspring express inbreeding depression. Immigrants arrive in the population and density regulation limits population growth. Key individual traits, parameter values, and variables are described in Table 1.

@@

After costs are paid, each remaining female chooses $N_{males}$ males to mate with, where $N_{males}$ is calculated by sampling from a Poisson distribution such that $N_{males} = Poisson(P_{p}) + 1$. This ensures that all surviving females choose at least one mate, and causes some stochastic variation around the expected mean $N_{males}$ of $P_{p}+1$. A female's value of $P_{p}$ depends on the sum of her 20 $P_{a}$ allele values, which might be negative or positive. However, because females cannot mate with a negative number of males, $P_{p}$ is constrained to be zero or positive such that $P_{p} = 0$ when $P_{g} < 0$, but $P_{p} = P_{g}$ when $P_{g} \geq 0$. Polyandry is therefore modelled as a threshold trait \cite[][]{Lynch1998, Roff1996, Roff1998, Duthie} that is influenced by continuous genetic variation but only expressed at the threshold value of $P_{g} > 0$. Any negative $P_{g}$ value therefore generates phenotypic monandry, while positive $P_{g}$ values can lead to different degrees of realised polyandry (i.e., different numbers of mates, $N_{males}$).

@@

The viability of a focal female $i$'s offspring ($\Psi_{\textrm{off}}$) decreases as a log-linear function of her kinship with the sire of her offspring $j$ ($k_{ij}$) and inbreeding depression slope $\beta$,
\begin{equation}
\Psi_{\textrm{off}} = e^{-\beta k_{ij}}
\end{equation}
Here, $\beta$ models the number of haploid lethal equivalents that exist as deleterious recessive alleles in the gametes of $i$ and $j$, and which might be homozygous in offspring and reduce viability. Equation 1 assumes independent allelic effects, generating multiplicative effects on offspring viability \cite[][]{Morton1956, Mills1994}. We model inbreeding depression as having an absolute rather than relative effect on offspring viability (i.e., hard rather than soft selection), so that the effect of $\beta$ is consistent across generations and different parameter combinations. We also assume that inbreeding always decreases offspring viability, $\beta > 0$ (i.e., no outbreeding depression). Therefore, because $k_{ij}$ is constrained to values between zero and one, $-\beta \times k_{ij} \leq 0$. Values of $\Psi_{\textrm{off}}$ must therefore be between zero (if $-\beta \times k_{ij}$ is very negative) and one (if $-\beta \times k_{ij} = 0$). We therefore define $\Psi_{\textrm{off}}$ as the probability that an offspring is viable, and sample its realised viability (versus mortality) using a Bernoulli trial.

@@

\subsection*{Simulation and analysis}

@@

When post-copulatory inbreeding avoidance alleles ($F_{a}$) were fixed to zero and $F_{p}$ could not evolve, $P_{a}$ values were always negative (Figure 1A,C,E,G). This shows that there is selection against unconditional polyandry even given zero direct cost ($c_{P}=0$; Figure 1A). This is because polyandrous females sampled more males from the available population and were consequently likely to mate with some relatives and hence produce some inbred offspring with low viability \cite[][]{Duthie}.

When post-copulatory inbreeding avoidance was allowed to evolve, mean $P_{a}$ values were substantially higher than in comparable simulations where post-copulatory inbreeding avoidance could not evolve (Figure 1B,D,F,H). Allowing evolution of post-copulatory inbreeding avoidance alongside pre-copulatory inbreeding avoidance therefore facilitated evolution of polyandry. However, $P_{a}$ values were expected to be positive only given a low cost of polyandry ($c_{P} < 0.005$, Figure 1B,D). This implies that, even given strong inbreeding depression in offspring viability, inbreeding avoidance might generally be a weak force driving polyandry evolution.

Strong post-copulatory inbreeding avoidance consistently evolved in all simulations where such evolution was allowed (Figure 1B,D,F,H), even when $P_{a}$ values were expected to be slightly negative (i.e., when there is selection against polyandry, Figure 1F,H). This reflects the threshold nature of phenotypic expression of polygenic polyandry, wherein random sampling of alleles means that polyandry is commonly expressed (i.e., $P_{g}>0$) even when mean $P_{a}$ values are negative (Figure 2). This means that there is commonly opportunity for expression of post-copulatory inbreeding avoidance and associated selection that drives evolution of $F_{a}$ allele values.

@@

 {\color{Gray}
\begin{adjustwidth}{-2.2in}{0in}
\mysidecaption{0.68}
{%
   \includegraphics[height=18cm]{figures/poly_costs.pdf}%
}%
{%
   \captionof{figures/poly_costs.pdf}\begin{justify}\vspace{10 mm} \textbf{Figure 1:} Mean allele values underlying tendency for polyandry (red), pre-copulatory inbreeding strategy (blue), and post-copulatory inbreeding strategy (black) from simulations where post-copulatory inbreeding strategy is (A, C, E, and G) fixed to zero (i.e., random fertilisation) or (B, D, F, H) allowed to evolve freely. Costs of polyandry ($c_{p}$) increase across rows from zero (A and B) to 0.01 (G and H). Mean allele values (solid lines) and associated standard errors (shading) are calculated across all individuals within a population over 40000 generations across 40 replicate populations. Negative mean allele values indicate inbreeding avoidance or tendency for monandry, and positive values indicate inbreeding preference or tendency for polyandry. Dotted lines demarcate mean allele values of zero. \end{justify}{\t}%
}[t]
\end{adjustwidth}
}

@@

The opportunity to adjust inbreeding has been widely considered as a driver of adaptive evolution of polyandry \cite[][]{Tregenza2002, Foerster2003, Akcay2007, Varian-Ramos2012, Kingma2013, Lehtonen2015, Reid2014}. Recently, \cite{Duthie} found that the conditions under which selection for polyandry and pre-copulatory inbreeding avoidance is predicted are highly restrictive. However, \cite{Duthie} assumed that fertilisation of offspring was unbiased with respect to mate kinship, meaning that females could not express post-copulatory choice. In contrast, we found that when post-copulatory inbreeding avoidance could evolve, selection for polyandry was greatly strengthened (Figure 1). The proposition that polyandry might facilitate cryptic female choice among males of varying compatibility is not new \cite[e.g.,][]{Zeh1997}, but the existence of selection for post-copulatory inbreeding avoidance causing feedback to select for polyandry has not been formally modelled. Our model therefore has widespread implications for future studies of evolution of polyandry.

@@

Recent studies have shown how females can successfully bias fertilisation after mating has occurred. For example, when female guppies (\textit{Poecilia reticulata}) were artificially inseminated with equal quantities of sperm from full-siblings and unrelated males, more eggs were fertilised by unrelated males because the velocities of full siblings' sperm were reduced by females' ovarian fluids \cite[][]{Gasparini2011}. In black field crickets (\textit{Teleogryllus commodus}), females attempt to remove the spermatophores of unwanted males after copulation, and are capable of controlling sperm transfer to spermatheca after copulation occurs \cite[][]{Bussiere2006, Tuni2013}. Our results suggest that evolution of such post-copulatory inbreeding avoidance might increase selection for further polyandry, especially if pre-copulatory inbreeding avoidance is costly (Figure 1).

@@

In our model, one cost free inbreeding strategy precluded another more costly inbreeding strategy from persisting in a focal population (Figure 3), meaning that long-term persistence of both pre-copulatory and post-copulatory inbreeding avoidance might not be expected in populations given cost asymmetry. Direct costs therefore strongly affected the evolution of the costly phenotype(s), meaning that quantifying costs is necessary for predicting evolution of phenotypes associated with inbreeding strategies. Quantifying direct costs of phenotypes associated with polyandry and mate choice is empirically challenging \cite[][]{Pomiankowski1987, Kokko2003, Reid2015}. \cite{Pomiankowski1987} categorised four types of costs that are relevant to mate choice, which we suggest are also relevant costs for polyandry, and include costs of elevated risks associated with (1) predation or (2) disease transmission, and costs incurred through (3) loss of time or (4) depletion of energy. In our model, we interpreted a cost of polyandy as an elevated risk of predation as might be incurred while searching for or courting mates \cite[e.g.,][]{Rowe1988, Rowe1994}, a cost of pre-copulatory inbreeding strategy as a lost of time \cite[i.e., risk of not finding a mate in time due to choosiness; e.g.,][]{Kokko2013}, and a cost of post-copulatory inbreeding strategy as depletion of energy. While these costs are biologically realistic, different costs will be relevant for different populations and be realised in different ways. For example, increased risk of disease transmission has been observed for highly polyandrous females \cite[][]{Roberts2015a}, but such a cost would more realistically be applied to the number of realised mates a female has rather than her tendency for engaging in polyandry. Future studies will therefore need to carefully consider the magnitudes and realisation of direct costs to accurately predict evolution of inbreeding strategies in populations.
